Article Title

### 2014-05-03 plos.tex
Response to reviewers.

\section*{Introduction}
\emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells. Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity. Although functional connectivity is a phenomenological description with no simple mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.

@@

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.  The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
\begin{equation}\label{eq:sample}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent). For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus. In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.

@@

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
    {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by green spheres and latent units by light-shaded spheres.  Edges represent non-zero partial correlations, \ie \sq{interactions}.
        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units. 
        {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units. \\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators. Some exceptions occur when the sample is too small to reveal the structure in the data. In these cases estimators with simpler structures can be more efficient than estimators with true structure.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{02}{Row 1, A}).  If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.

The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$. The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{02}{Row 1, B}).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).

Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix. The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{02}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.

@@

\paragraph{Simulation}
We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{02}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).  We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes. Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{02}{\,Row 3}).

@@

This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$,
\begin{equation}
     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
\end{equation}
Normalization by $p$ makes the values of the loss function comparable across different population sizes.

Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.

@@

We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the excess loss for each of the five estimators.  The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.  However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).  There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).  Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because it reliably inferred the correct number of latent units: zero.

@@

With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly. Instead, the loss function must be estimated from the sample through \emph{validation}.  In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.   The estimate $C_{\sf 0}^\prime$ is used to 
compute the \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.

The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument, \ie
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}
With this property of additivity, the validation loss is an unbiased estimate of the loss function:
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}.
\end{equation*}

@@

Acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as the training sample. The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.  Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ when the remaining $K-1$ subsets are treated as the training sample. Then the cross-validation loss for estimator $C$ is
\begin{equation}\label{eq:cv-loss}
    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
\end{equation}
In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier). However, the present framework is more general, as other loss functions can be used.

Since we only need to compare estimators to each other, we are  interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$,
\begin{equation}\label{eq:rel-cv-loss}
    \Delta\ell_{C,C_{\sf ref}} = \ell_{C}-\ell_{C_{\sf ref}}
\end{equation}
The relative cross-validation loss was used for the main findings in this study.

To examine how well the relative CV-losses reproduced the relationships between the excess losses, we evaluated the relative CV-loss with respect to the estimator that matched the true structure of the model. Indeed, CV-losses accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{02}{Row 6}).

@@

\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{03}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}. This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}).  The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).  First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.  Each grating was presented for 500 ms, without intervening blanks.  This stimulus was used to compute the orientation tuning of the recorded cells (\figref{03}{E}). To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with between 100 and 300 repetitions of each direction. In this stimulus, each grating lasted 1 second and was followed by a 1-second blank.  The average stimulus response was subtracted from each trial. The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.   The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{under 0.04 in all sites}) (\figref{03}{G} and \figref{06}{D}).

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}). 
    {\bf C--G.} Calcium activity data from an example site.
    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.  The traces from {\bf C} are highlighted in red.
    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site. Only spiking cells are shown. The cells' colors indicate their orientation preferences. The gray cells were not significantly tuned.
    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
} \label{fig:03}}
    {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
\end{figure}

@@

The same evaluation based on the quadratic loss function,
\begin{equation}\label{eq:quadratic}
\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2,
\end{equation}
instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S\ref{supp:02}). This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.

@@

We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:05} and Fig.~\ref{fig:06}). Although the regularized estimates were similar to the sample correlation matrix (\figref{05}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{05}{B and E}). The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{05}{C}).

>>>>

\section*{Introduction}
\emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells. Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity. Although functional connectivity is a phenomenological description without a strict mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.

@@

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf sample}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \ie toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation of the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.  The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf sample}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
\begin{equation}\label{eq:sample}
    C_{\sf sample} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent). For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus. In this study, we estimate the true mean, $\mu$, using the sample mean, $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf sample}$.

@@

To see that this is the case, partition the covariance matrix 
\begin{equation}
C = \begin{pmatrix}
C_{11} & C_{12} \\
C_{12}^\T & C_{22}
\end{pmatrix}
\end{equation}
so that $C_{11}$ is the $2\times 2$ covariance matrix of two cells of interest. Such partitioning can be done for any pair of cells by changing their order. 
Then the covariance matrix of the conditional distribution of these two cells is \cite{Anderson:2003}:
\begin{equation}
\hat C_{11} = C_{11}-C_{12}C^{-1}C_{12}^\T 
\end{equation}
The inverse of the conditional covariance matrix can be found in the corresponding partition of the precision matrix $C^{-1}$:
\begin{equation}
C^{-1}=\begin{pmatrix}
\hat C_{11}^{-1} &   -\hat C_{11}^{-1}C_{12}C_{22}^{-1} \\
-C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}  &  C_{22}^{-1}+C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}C_{12}C_{22}^{-1}
\end{pmatrix}
\end{equation}
The inverse of a $2\times 2$ correlation matrix reverses the sign of its off-diagonal elements:
\begin{equation}
\begin{pmatrix}
1 & \rho \\
\rho & 1
\end{pmatrix}^{-1}
\propto 
\begin{pmatrix}
-1 & \rho \\
\rho & -1 
\end{pmatrix}
\end{equation}
Therefore, scaling $C^{-1}$ to have 1s on the diagonal yields the partial correlation in each of its $2\times 2$ partitions, resulting in a matrix of partial pairwise correlations.

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the target estimates of these estimators correspond to distinct families of graphical models (\figref{1}{\,Row 1}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
    {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by green spheres and latent units by light-shaded spheres.  Edges represent non-zero partial correlations, \ie \sq{interactions}.
        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units. 
        {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units. \\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators. Some exceptions occur when the sample is too small to reveal the structure in the data. In these cases estimators with simpler structures can be more efficient than estimators with true structure.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:1}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf sample}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf sample} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{1}{Row 1, A}).  If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.

The target of estimator $C_{\sf factor}$ is the factor model $L + D$, where $L$ is the $p\times p$ positive \hl{semidefinite} matrix with low rank and $D$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf sample} + \lambda (L + D),
\end{equation}
has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$. The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between the recorded cells are insignificant (\figref{1}{Row 1, B}).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{1}{Row 1, C}).

Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix. The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{1}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.

@@

\paragraph{Simulation}
We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{1}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).  We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes. Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{1}{\,Row 3}).

@@

This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf sample})$,
\begin{equation}
     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
\end{equation}
Normalization by $p$ makes the values of the loss function comparable across different population sizes.

Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf sample})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf sample}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.

@@

We drew 30 independent samples of sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the excess loss for each of the five estimators.  The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.  However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{1}{\,row 5}).  There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{1}{\,Row 5, C and D}).  Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because it reliably inferred the correct number of latent units: zero.

@@

With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly. Instead, the loss function must be estimated from the sample through \emph{validation}.  In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf sample}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.   The estimate $C_{\sf sample}^\prime$ is used to 
compute the \emph{validation loss} $\loss{C,C_{\sf sample}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.

The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument, \ie
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}
With this property of additivity, the validation loss is an unbiased estimate of the loss function:
\begin{equation*}
    \E[C_{\sf sample}^\prime]{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E[C_{\sf sample}^\prime]{C_{\sf sample}^\prime}}=\loss{C,\Sigma}.
\end{equation*}

@@

Acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as the training sample. The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss}.

Cross-validation loss accurately reproduced the relationships between the excess losses of the estimators, with somewhat greater variability (\figref{1}{Row 6}).

@@

\paragraph{The $C_{\sf sparse+latent}$ estimator is most efficient in neural data}
We recorded the calcium activity of dense populations of neurons in the supragranular layers in primary visual cortex of fentanyl-anesthetized mice using fast random-access 3D scanning two-photon microscopy during visual stimulation (\figref{2}{A--B}) \cite{Reddy:2005, Katona:2012, Cotton:2013}. This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{2}{C}).  The firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{2}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (For details, see Methods).  First, 30 repetitions of full-field drifting gratings of 16 directions were presented in random order.  Each grating was presented for 500 ms, without intervening blanks.  This stimulus was used to compute the orientation tuning of the recorded cells (\figref{2}{E}). To estimate the noise correlation matrix, we presented only 2 distinct directions in some experiments or 5 directions in others with between 100 and 300 repetitions of each direction. In this stimulus, each grating lasted 1 second and was followed by a 1-second blank.  The average stimulus response was subtracted from each trial. The traces were then binned into 150 ms intervals for the estimation of the covariance matrix.   The sample correlation coefficients were largely positive and low (average across sites of 0.02 and \hl{less than 0.04 in any site}) (\figref{2}{G} and \figref{5}{D}).

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}). 
    {\bf C--G.} Calcium activity data from an example site.
    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.  The traces from {\bf C} are highlighted in red.
    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site. Only spiking cells are shown. The cells' colors indicate their orientation preferences. The gray cells were not significantly tuned.
    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
} \label{fig:2}}
    {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
\end{figure}

@@

The same evaluation based on the quadratic loss function,
\begin{equation}\label{eq:quadratic}
\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
\end{equation}
instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S\ref{supp:02}). This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.

@@

We examined the composition of the $C_{\sf sparse+latent}$ estimates at each imaged site (Fig.~\ref{fig:4} and Fig.~\ref{fig:5}). Although the regularized estimates were similar to the sample correlation matrix (\figref{4}{A and D}), the corresponding partial correlation matrices differed substantially (\figref{4}{B and E}). The partial correlation matrix of the regularized estimate was the sum of a sparse component and a low-rank component (\figref{4}{C}).

In the example site (Fig.~\ref{fig:4}), the sparse component had 89.9\% sparsity (or conversely, 11.1\% connectivity: $connectivity=1-sparsity$) with average node degree of 29.3 (\figref{4}{G}). The average node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons. The low-rank component had rank 71, denoting 71 inferred latent units. The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{5}{A and B}). There was an inverse relationship between the number of latent units and the average node degree (\figref{5}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions. This variability may be explained by differences in brain states and recording quality and needs further investigation.

The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 31 sites were about 5 times lower than the average sample correlations (\figref{5}{D}). This suggests that correlations between neurons build up from large numbers of small interactions with other neurons. Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.41 whereas that of the average partial correlations was 0.35.
